### 2025-03-12 code/Reddit_eda.qmd
adding eda code

---
title: "Assignment 2"
author: Group 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: false
  tidy: true
format: pdf
theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(fedmatch)
library(tidyverse)
library(readxl)
```

# We will explore the data we scraped to determine next steps for analysis.

1.  Read in data from github, and tidy it to continue processing.

```{r}
# posts data
reddit_posts <- read_xlsx(
  "~/repos/SURV622_Assignment/data/posts_data_clean.xlsx") |> 
  mutate(date_utc = ymd(date_utc)) |> 
  # remove variables not needed
  select(-timestamp)

# comments data
reddit_comment <- read_xlsx(
  "~/repos/SURV622_Assignment/data/comments_data_clean.xlsx") |> 
    mutate(date_utc = ymd(date_utc)) |> 
  # great unique post id
  arrange(date_utc) |> 
  group_by(url) |> 
  mutate(post_id = cur_group_id()) |> 
  ungroup() |> 
  # remove variables not needed
  select(-author, -timestamp, -comment_id)

# id url group IDs to the posts data
reddit_posts_2 <- reddit_posts |> 
  semi_join(reddit_comment |> select(url, post_id))

```

### The strings we will be processing sometimes have special characters, numbers, brackets. We need to clean these string for processing.

-   We can use the fedmatch::clean_strings() function to tidy up the text. The following is an example of what will be cleaned.

```{r}
# print example of posts
reddit_posts |> 
  slice(10:15) |> 
  select(text)

# clean posts
reddit_posts |> 
  slice(10:15) |> 
  mutate(text = clean_strings(text)) |> 
  select(text)

```

-   Clean posts and comment strings.

```{r}
reddit_posts <- reddit_posts |> 
  mutate(text = clean_strings(text))

reddit_comment <- reddit_comment |> 
  mutate(comment = clean_strings(comment))

reddit_posts_2 <- reddit_posts_2 |> 
  mutate(text=clean_strings(text))
```

\

### We can now begin to explore the data. First we explore the posts.

```{r, results = 'asis'}

  reddit_posts |>
  slice(364) |> 
  pull(text)

```