### 2025-03-13 code/Reddit_sentiment.qmd
adding sentiment analysis

---
title: "Assignment 2"
author: Group 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: false
  tidy: true
format: pdf
theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(writexl)
library(text)
library(readxl)
library(tidyverse)

```

We will explore sentiment analysis to give us an advantage in coding comments.

Read in data from github, and tidy it to continue processing.

```{r}

reddit_posts <- read_xlsx(
  "~/repos/repos/SURV622_Assignment/data/posts_data_clean.xlsx")

# comments data
reddit_comment <- read_xlsx(
  "~/repos/repos/SURV622_Assignment/data/comments_data_clean.xlsx")

```

We can use a model from huggingface to code each comment as positive, negative, or neutral.

We build a function that we will use to loop through comments and get scores.

```{r}
sentiment_score_fun <- function(comment_text){

  description = str_trunc(comment_text, 1600)

  # shorten descriptions to confirm to the model limitation

  # use huggingface model to detect emotional tone of summary
  mod = 
    textClassify(description,
                 model_info = "ProsusAI/finbert", 
                 return_all_scores = FALSE, 
                 return_incorrect_results = FALSE,
                 function_to_apply = "softmax",
                 tokenizer_parallelism = TRUE,
                 device = "gpu") |>   # change to "cpu" if a gpu is not configured
    pull(1)

  return(mod)

}
```

Loop the comments through our function.

```{r}
reddit_comment_new <- reddit_comment |> 
  mutate(label = ifelse(is.na(comment), NA, sentiment_score_fun(comment)))

```

```{r}
 reddit_comment_new |> count(label)
```

```{r}
 reddit_posts_new <- reddit_posts |> mutate(label = ifelse(is.na(text), NA, sentiment_score_fun(text)))
```

```{r}
 reddit_posts_new |> count(label)
```

```{r}
# write posts data
writexl::write_xlsx(reddit_posts_new, "~/repos/repos/SURV622_Assignment/data/posts_data_clean.xlsx")

# write comments data
writexl::write_xlsx(reddit_comment_new, "~/repos//repos/SURV622_Assignment/data/comments_data_clean.xlsx")

```