### 2025-03-13 code/Reddit_eda.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Assignment 2},
  pdfauthor={Group 2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Assignment 2}
\author{Group 2}
\date{2025-03-13}

\begin{document}
\maketitle

\section{We will explore the data we scraped to determine next steps for
analysis.}\label{we-will-explore-the-data-we-scraped-to-determine-next-steps-for-analysis.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Read in data from github, and tidy it to continue processing.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# posts data}
\NormalTok{reddit\_posts }\OtherTok{\textless{}{-}} \FunctionTok{read\_xlsx}\NormalTok{(}
  \CommentTok{\# "\textasciitilde{}/repos/SURV622\_Assignment/data/posts\_data\_clean.xlsx") |\textgreater{} path for Kevin}
  \StringTok{"\textasciitilde{}/Desktop/UMICH/SURV622\_Assignment/data/posts\_data\_clean.xlsx"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \CommentTok{\# path for Felix}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date\_utc =} \FunctionTok{ymd}\NormalTok{(date\_utc)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# remove variables not needed}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{timestamp)}

\CommentTok{\# comments data}
\NormalTok{reddit\_comment }\OtherTok{\textless{}{-}} \FunctionTok{read\_xlsx}\NormalTok{(}
  \CommentTok{\# "\textasciitilde{}/repos/SURV622\_Assignment/data/comments\_data\_clean.xlsx") |\textgreater{} Kevin\textquotesingle{}s path}
  \StringTok{"\textasciitilde{}/Desktop/UMICH/SURV622\_Assignment/data/comments\_data\_clean.xlsx"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} \CommentTok{\# Felix\textquotesingle{}s path}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date\_utc =} \FunctionTok{ymd}\NormalTok{(date\_utc)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# great unique post id}
  \FunctionTok{arrange}\NormalTok{(date\_utc) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{group\_by}\NormalTok{(url) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{post\_id =} \FunctionTok{cur\_group\_id}\NormalTok{()) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ungroup}\NormalTok{() }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# remove variables not needed}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{author, }\SpecialCharTok{{-}}\NormalTok{timestamp, }\SpecialCharTok{{-}}\NormalTok{comment\_id) }

\CommentTok{\# id url group IDs to the posts data}
\NormalTok{reddit\_posts\_2 }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{semi\_join}\NormalTok{(reddit\_comment }\SpecialCharTok{|\textgreater{}} \FunctionTok{select}\NormalTok{(url, post\_id)) }
\end{Highlighting}
\end{Shaded}

\subsubsection{The strings we will be processing sometimes have special
characters, numbers, brackets. We need to clean these string for
processing.}\label{the-strings-we-will-be-processing-sometimes-have-special-characters-numbers-brackets.-we-need-to-clean-these-string-for-processing.}

\begin{itemize}
\tightlist
\item
  We can use the fedmatch::clean\_strings() function to tidy up the
  text. The following is an example of what will be cleaned.
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# print example of posts}
\NormalTok{reddit\_posts }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{slice}\NormalTok{(}\DecValTok{10}\SpecialCharTok{:}\DecValTok{15}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{select}\NormalTok{(text) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 6 x 1
  text                                                                          
  <chr>                                                                         
1 <NA>                                                                          
2 <NA>                                                                          
3 <NA>                                                                          
4 <NA>                                                                          
5 <NA>                                                                          
6 commerce secretary howard lutnick said sunday that government spending could ~
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# clean posts}
\NormalTok{reddit\_posts }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{slice}\NormalTok{(}\DecValTok{10}\SpecialCharTok{:}\DecValTok{15}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{text =} \FunctionTok{clean\_strings}\NormalTok{(text)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{select}\NormalTok{(text) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 6 x 1
  text                                                                          
  <chr>                                                                         
1 <NA>                                                                          
2 <NA>                                                                          
3 <NA>                                                                          
4 <NA>                                                                          
5 <NA>                                                                          
6 commerce secretary howard lutnick said sunday that government spending could ~
\end{verbatim}

\begin{itemize}
\tightlist
\item
  Clean posts and comment strings.
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{reddit\_posts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{text =} \FunctionTok{clean\_strings}\NormalTok{(text))}

\NormalTok{reddit\_comment }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_comment }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{comment =} \FunctionTok{clean\_strings}\NormalTok{(comment))}

\NormalTok{reddit\_posts\_2 }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts\_2 }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{text=}\FunctionTok{clean\_strings}\NormalTok{(text))}
\end{Highlighting}
\end{Shaded}

\hfill\break

\subsubsection{We can now begin to explore the data. First we explore
the
posts.}\label{we-can-now-begin-to-explore-the-data.-first-we-explore-the-posts.}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{  reddit\_posts }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{slice}\NormalTok{(}\DecValTok{364}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{pull}\NormalTok{(text) }
\end{Highlighting}
\end{Shaded}

{[}1{]} ``hello all so i recently took the deferred resignation i was
working at the va and my job was not exempt i did not trust that my job
was safe even after coming off probation the day the fork in the road
came out i took the resignation and my last day was february 28 before i
started i was in the cwt program compensated work therapy i once again
have an opportunity to get back into the program i know we are allowed
to work while on the resignation but am i able to do the cwt because
both checks would be federal payments i don t want to lose out on the
resignation benefits just to make 12 hr but working does help me mental
health and give me a purpose any insight would be greatly appreciated''

How many posts were collected in total and by day? How many posts after
data cleaning? Is there a pattern with respect to the time of day or day
of week when posts were created?

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# number of total posts}
\NormalTok{totalposts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize}\NormalTok{(}
    \AttributeTok{TotalPosts =} \FunctionTok{n}\NormalTok{()}
\NormalTok{  )}
\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(totalposts)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}r@{}}
\toprule\noalign{}
TotalPosts \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
557 \\
\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# number of posts by day}
\NormalTok{byday }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(date\_utc) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize}\NormalTok{(}
    \AttributeTok{total =} \FunctionTok{n}\NormalTok{()}
\NormalTok{  )}

\CommentTok{\# plot}
\FunctionTok{ggplot}\NormalTok{(byday, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ date\_utc, }\AttributeTok{y =}\NormalTok{ total)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{size =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+}  \CommentTok{\# Larger points for visibility}
  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{linewidth =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{+}  \CommentTok{\# Thicker line for clarity}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{title =} \StringTok{"Total Number of Posts per Day"}\NormalTok{,}
    \AttributeTok{x =} \StringTok{"Day"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Total Number of Posts"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Reddit_eda_files/figure-pdf/unnamed-chunk-5-1.pdf}

Is there a relationship between events and frequency of tweets?

The two peaks in our graph can potentially be attributed to these two
major news stories on those days. On March 5, Elon talked about how he
wishes to ``save Western civilization from empathy''. On March 6,
President Trump was reported to further limit Elon Musk's authority amid
backlash to DOGE cuts. These two events may have had a major impact in
the number of Reddit posts related to DOGE and Elon.

What are the words in the set of posts you have assembled that appear
most frequently?

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# tokenize words to count frequencies for reddit posts}
\NormalTok{posts\_word\_counts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{select}\NormalTok{(title, text) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(title, text), }\AttributeTok{values\_drop\_na =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest\_tokens}\NormalTok{(word, value) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{count}\NormalTok{(word, }\AttributeTok{sort =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{slice\_max}\NormalTok{(n, }\AttributeTok{n =} \DecValTok{50}\NormalTok{)}

\CommentTok{\# plot}
\FunctionTok{ggplot}\NormalTok{(posts\_word\_counts, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{reorder}\NormalTok{(word, n), }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_col}\NormalTok{(}\AttributeTok{fill =} \StringTok{"steelblue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{coord\_flip}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Top 50 Words in Reddit Posts (Including Stopwords)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Word"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Reddit_eda_files/figure-pdf/unnamed-chunk-6-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# tokenize words to count frequencies for reddit comments}
\NormalTok{comments\_word\_counts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_comment }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{select}\NormalTok{(comment) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(comment), }\AttributeTok{values\_drop\_na =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest\_tokens}\NormalTok{(word, value) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{count}\NormalTok{(word, }\AttributeTok{sort =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{slice\_max}\NormalTok{(n, }\AttributeTok{n =} \DecValTok{50}\NormalTok{)}

\CommentTok{\# plot}
\FunctionTok{ggplot}\NormalTok{(comments\_word\_counts, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{reorder}\NormalTok{(word, n), }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_col}\NormalTok{(}\AttributeTok{fill =} \StringTok{"lightgreen"}\NormalTok{) }\SpecialCharTok{+}  \CommentTok{\# Removed extra parentheses}
  \FunctionTok{coord\_flip}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Top 50 Words in Reddit Comments (Including Stopwords)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Word"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Reddit_eda_files/figure-pdf/unnamed-chunk-6-2.pdf}

The above graphs show the most frequent words in both reddit posts and
their comments. Most of these are stopwords, therefore we will proceed
to remove these to get a better idea of what people are really
discussing.

How does this change if you exclude ``stop words'' such as ``a,''
``an,'' ``the,'' ``is,'' and others that are common in English sentences
but are generally not informative?

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# tokenize words to count frequencies for reddit posts removing stop words}
\NormalTok{posts\_word\_counts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_posts }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{select}\NormalTok{(title, text) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(title, text), }\AttributeTok{values\_drop\_na =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest\_tokens}\NormalTok{(word, value) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{anti\_join}\NormalTok{(stop\_words, }\AttributeTok{by =} \StringTok{"word"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{count}\NormalTok{(word, }\AttributeTok{sort =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{slice\_max}\NormalTok{(n, }\AttributeTok{n =} \DecValTok{50}\NormalTok{)}

\CommentTok{\# plot}
\FunctionTok{ggplot}\NormalTok{(posts\_word\_counts, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{reorder}\NormalTok{(word, n), }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_col}\NormalTok{(}\AttributeTok{fill =} \StringTok{"steelblue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{coord\_flip}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Top 50 Words in Reddit Posts (Removing Stopwords)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Word"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Reddit_eda_files/figure-pdf/unnamed-chunk-7-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# tokenize words to count frequencies for Reddit comments, removing stop words}
\NormalTok{comments\_word\_counts }\OtherTok{\textless{}{-}}\NormalTok{ reddit\_comment }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{select}\NormalTok{(comment) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(comment), }\AttributeTok{values\_drop\_na =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{unnest\_tokens}\NormalTok{(word, value) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{anti\_join}\NormalTok{(stop\_words, }\AttributeTok{by =} \StringTok{"word"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{count}\NormalTok{(word, }\AttributeTok{sort =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{slice\_max}\NormalTok{(}\AttributeTok{order\_by =}\NormalTok{ n, }\AttributeTok{n =} \DecValTok{50}\NormalTok{)}

\CommentTok{\# plot}
\FunctionTok{ggplot}\NormalTok{(comments\_word\_counts, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{reorder}\NormalTok{(word, n), }\AttributeTok{y =}\NormalTok{ n)) }\SpecialCharTok{+}
  \FunctionTok{geom\_col}\NormalTok{(}\AttributeTok{fill =} \StringTok{"lightgreen"}\NormalTok{) }\SpecialCharTok{+}  \CommentTok{\# Removed extra parentheses}
  \FunctionTok{coord\_flip}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Top 50 Words in Reddit Comments (Removing Stopwords)"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Word"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Reddit_eda_files/figure-pdf/unnamed-chunk-7-2.pdf}

The above graphs show the most frequent words used in both Reddit posts
and their comment sections. We see that most of these are directly
related to our topic, with mentions of the President, the federal
government, DOGE and RIF.

\end{document}