### 2025-03-14 code/Reddit_eda.qmd
Added interpretations to analysis graphs

---
title: "Assignment 2"
author: Group 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: false
  tidy: true
format: pdf
theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(fedmatch)
library(tidyverse)
library(readxl)
library(tidytext)
```

# We will explore the data we scraped to determine next steps for analysis.

1.  Read in data from github, and tidy it to continue processing.

```{r}
# posts data
reddit_posts <- read_xlsx(
  # "~/repos/SURV622_Assignment/data/posts_data_clean.xlsx") |> path for Kevin
  "~/Desktop/UMICH/SURV622_Assignment/data/posts_data_clean.xlsx") |> # path for Felix
  mutate(date_utc = ymd(date_utc)) |> 
  # remove variables not needed
  # select(-timestamp)
  mutate(
    # Convert timestamp to POSIXct format
    timestamp = as.POSIXct(timestamp, origin = "1970-01-01", tz = "UTC"),

    # Extract time of day (you can adjust the format as needed)
    time_of_day = format(timestamp, "%H:%M:%S")
  )

# comments data
reddit_comment <- read_xlsx(
  # "~/repos/SURV622_Assignment/data/comments_data_clean.xlsx") |> Kevin's path
  "~/Desktop/UMICH/SURV622_Assignment/data/comments_data_clean.xlsx") |> # Felix's path
    mutate(date_utc = ymd(date_utc)) |> 
  # great unique post id
  arrange(date_utc) |> 
  group_by(url) |> 
  mutate(post_id = cur_group_id()) |> 
  ungroup() |> 
  # remove variables not needed
  mutate(
    # Convert timestamp to POSIXct format
    timestamp = as.POSIXct(timestamp, origin = "1970-01-01", tz = "UTC"),

    # Extract time of day (you can adjust the format as needed)
    time_of_day = format(timestamp, "%H:%M:%S")
  ) |>
  select(-author, -comment_id)

# id url group IDs to the posts data
reddit_posts_2 <- reddit_posts |> 
  semi_join(reddit_comment |> select(url, post_id))

```

### The strings we will be processing sometimes have special characters, numbers, brackets. We need to clean these string for processing.

-   We can use the fedmatch::clean_strings() function to tidy up the text. The following is an example of what will be cleaned.

```{r}
# print example of posts
reddit_posts |> 
  slice(10:15) |> 
  select(text)

# clean posts
reddit_posts |> 
  slice(10:15) |> 
  mutate(text = clean_strings(text)) |> 
  select(text)

```

-   Clean posts and comment strings.

```{r}
reddit_posts <- reddit_posts |> 
  mutate(text = clean_strings(text))

reddit_comment <- reddit_comment |> 
  mutate(comment = clean_strings(comment))

reddit_posts_2 <- reddit_posts_2 |> 
  mutate(text=clean_strings(text))
```

\

### We can now begin to explore the data. First we explore the posts.

```{r, results = 'asis'}

  reddit_posts |>
  slice(364) |> 
  pull(text)

```

## How many posts were collected in total and by day? How many posts after data cleaning? Is there a pattern with respect to the time of day or day of week when posts were created? Is there a relationship between events and frequency of tweets?
```{r}
# number of total posts
totalposts <- reddit_posts %>%
  summarize(
    TotalPosts = n()
  )
knitr::kable(totalposts)

# number of posts by day
byday_posts <- reddit_posts %>%
  group_by(date_utc) %>%
  summarize(
    total = n()
  )

# plot
ggplot(byday_posts, aes(x = date_utc, y = total)) + 
  geom_point(color = "blue", size = 3) +  
  geom_line(color = "steelblue", linewidth = 1) +  
  labs(
    title = "Total Number of Posts per Day",
    x = "Day",
    y = "Total Number of Posts"
  ) +
  theme_minimal()

# number of total comments
totalcomments <- reddit_comment %>%
  summarize(
    TotalComments = n()
  )
knitr::kable(totalcomments)

# number of comments by day
byday_comments <- reddit_comment %>%
  group_by(date_utc) %>%
  summarize(
    total = n()
  )

# plot
ggplot(byday_comments, aes(x = date_utc, y = total)) + 
  geom_point(color = "green", size = 3) +  
  geom_line(color = "lightgreen", linewidth = 1) +
  labs(
    title = "Total Number of Posts per Day",
    x = "Day",
    y = "Total Number of Posts"
  ) +
  theme_minimal()
```

There appears to be a noticeable peak in activity during the middle of the week. However, we don't have enough data to conclusively say it's due to the day of the week itself. It's more likely that these spikes are linked to specific events that occurred on those days, which sparked conversations.

The two peaks in our graph could be attributed to two major news stories. On March 5, Elon Musk made a statement about wanting to "save Western civilization from empathy," and on March 6, there were reports that President Trump was limiting Musk's authority due to backlash over cuts to DOGE. These events likely had a significant impact on the volume of Reddit posts related to Elon Musk and DOGE, driving the observed spikes in activity.

```{r}
# Number of posts by time of day (hour)
byhour_posts <- reddit_posts %>%
  mutate(hour_of_day = hour(timestamp)) %>%
  group_by(hour_of_day) %>%
  summarize(total = n())

# Plot number of posts by hour of day
ggplot(byhour_posts, aes(x = hour_of_day, y = total)) + 
  # geom_bar(stat = "identity", fill = "steelblue") + 
  geom_point(color = "blue", size = 3) +  
  geom_line(color = "steelblue", linewidth = 1) +
  labs(
    title = "Total Number of Posts by Hour of Day (24-hour clock)",
    x = "Hour of Day",
    y = "Total Number of Posts"
  ) +
  theme_minimal()

# Number of comments by time of day (hour)
byhour_comments <- reddit_comment %>%
  mutate(hour_of_day = hour(timestamp)) %>%
  group_by(hour_of_day) %>%
  summarize(total = n())

# Plot number of posts by hour of day
ggplot(byhour_comments, aes(x = hour_of_day, y = total)) + 
  # geom_bar(stat = "identity", fill = "lightgreen") + 
  geom_point(color = "green", size = 3) +  
  geom_line(color = "lightgreen", linewidth = 1) +
  labs(
    title = "Total Number of Comments by Hour of Day (24-hour clock)",
    x = "Hour of Day",
    y = "Total Number of Posts"
  ) +
  theme_minimal()
```

Most posts occur during the early morning, noon, and afternoon hours. The highest peak is around lunchtime, likely due to people posting during their lunch break. There is also significant activity in the afternoon, evening, and early morning. This could be because people have more free time or stay up late, allowing them to take the time to write posts, which require more effort than quick interactions. The fewest posts are seen in the morning, which makes sense as people are typically getting ready for work, commuting, or sleeping in.

The trend for comments mirrors that of posts. This makes sense because people are likely using Reddit at similar times for both posting and commenting. However, the volume of comments is much higher than the number of posts, as each post receives many comments from different users.

## What are the words in the set of posts you have assembled that appear most frequently? 
```{r}
# tokenize words to count frequencies for reddit posts
posts_word_counts <- reddit_posts |>
  select(title, text) |>
  pivot_longer(cols = c(title, text), values_drop_na = TRUE) |>
  unnest_tokens(word, value) |>
  count(word, sort = TRUE) |>
  slice_max(n, n = 50)

# plot
ggplot(posts_word_counts, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 50 Words in Reddit Posts (Including Stopwords)",
       x = "Word",
       y = "Count") +
  theme_minimal()

# tokenize words to count frequencies for reddit comments
comments_word_counts <- reddit_comment |>
  select(comment) |>
  pivot_longer(cols = c(comment), values_drop_na = TRUE) |>
  unnest_tokens(word, value) |>
  count(word, sort = TRUE) |>
  slice_max(n, n = 50)

# plot
ggplot(comments_word_counts, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "lightgreen") +  # Removed extra parentheses
  coord_flip() +
  labs(title = "Top 50 Words in Reddit Comments (Including Stopwords)",
       x = "Word",
       y = "Count") +
  theme_minimal()
```

The graphs above show the most frequent words in both Reddit posts and their comments. However, many of these words are stopwords, which don't provide much insight into the actual content of the discussions. To get a better understanding of what people are really talking about, we will remove these stopwords in the next step.

## How does this change if you exclude “stop words” such as “a,” “an,” “the,” “is,” and others that are common in English sentences but are generally not informative?
```{r}
# tokenize words to count frequencies for reddit posts removing stop words
posts_word_counts <- reddit_posts |>
  select(title, text) |>
  pivot_longer(cols = c(title, text), values_drop_na = TRUE) |>
  unnest_tokens(word, value) |>
  anti_join(stop_words, by = "word") |>
  count(word, sort = TRUE) |>
  slice_max(n, n = 50)

# plot
ggplot(posts_word_counts, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 50 Words in Reddit Posts (Removing Stopwords)",
       x = "Word",
       y = "Count") +
  theme_minimal()

# tokenize words to count frequencies for Reddit comments, removing stop words
comments_word_counts <- reddit_comment |>
  select(comment) |>
  pivot_longer(cols = c(comment), values_drop_na = TRUE) |>
  unnest_tokens(word, value) |>
  anti_join(stop_words, by = "word") |>
  count(word, sort = TRUE) |>
  slice_max(order_by = n, n = 50)

# plot
ggplot(comments_word_counts, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "lightgreen") +  # Removed extra parentheses
  coord_flip() +
  labs(title = "Top 50 Words in Reddit Comments (Removing Stopwords)",
       x = "Word",
       y = "Count") +
  theme_minimal()
```

The graphs above show the most frequent words used in both Reddit posts and their comment sections. We can see that many of these words are directly related to our topic, with frequent mentions of the President, DOGE, the federal government, and RIF.