### 2025-03-27 code/ml_llm/R/MLtexttutorial.Rmd
assignment 3 materials

---
title: "Machine Learning for Automated Stance Detection"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE)
```

This notebook gives a simple example of using supervised machine learning to automatically detect stance of topic from tweets. There's a lot of code here -- feel free to copy and paste much of it as you work on the exercises and you do your assignment. This is also only barely scratching the surface, and intended to be as transparent of an introduction as possible in terms of showing how everything is done. Of note, the `caret` package in R can be extremely helpful in doing some of what's done here with much less code involved (see https://www.machinelearningplus.com/machine-learning/caret-package/ for more information). However, there's somewhat of a learning curve associated with it, as well as the inability to see how everything works, so I only use the `caret` package to do a few things, like calculating precision and recall.

```{r}
# If you have not installed these, make sure to install first!

# Tidyverse and text analysis tools
library(tidyverse)
library(tidytext)

# Package for text mining
library(tm)

# Package for Stemming and Lemmatizing 
library(textstem)

# For SVM
library(e1071)

# For nearest neighbors
library(class)

# For ML evaluation
library(caret)
```

# Data

We downloaded the data from SemEval2016 website, and the details of the dataset can be seen in [SemEval2016 Task 6](https://aclanthology.org/S16-1003/).

```{r}
semeval_stance <- read.csv('../../data/SemEval2016-testdata-taskA-all-annotations.csv', stringsAsFactors = FALSE,)
semeval_stance$ID <- as.character(semeval_stance$ID) # Need to make this character for later joins
semeval_stance$Stance <- as.factor(semeval_stance$Stance) #Convert Char to Factor
# Also we need to remove #SemST from the Tweet column
semeval_stance$Tweet <- gsub("#SemST", "", semeval_stance$Tweet)
glimpse(semeval_stance)
```

Here, we have a few key variables. The `ID` variable is simply numbers from 1 to 4036, denoting the unique tweet. The tweet's `Target` and `Tweet` (i.e., the content) are all included. Finally, the `Stance` denotes what opinion the tweets express, Favore, Oppose, or . Our goal is to build a machine learning model that takes this data and is able to predict the topic of the tweet based on the content of the tweet.

We want to do the following steps to turn this text data into features:

-**Tokenize:** Split up the tweets into individual words

-**Stop Words:** Remove words that are too frequent and uninformative, like "a", "an", and "the".

-**Bag of Words:** We want columns representing all words in the entire corpus, and the rows representing tweets, with each cell indicating the counts of words in that tweet. This matrix is also known as a term-document matrix.

We'll look at the distribution of words as we go to inform how we're doing. Let's first tokenize and then look at what words in our corpus. First, let's tokenize and count the number of instances of each word. The key function we're using here is `unnest_tokens`, which we're using the break up the big tweet string for each tweet into individual strings with just one word each.

```{r}
semeval_stance %>% 
  unnest_tokens(word, 'Tweet') %>%  # tokenize
  count(word, sort = TRUE) %>% # count by word
  arrange(desc(n)) %>% # Everything from this point on is just to graph
  head(20) %>%
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_bar(stat= 'identity') + 
  coord_flip()
```

Looks like there are a lot of stop words being caught! Let's take those out. To do this, we use the `stop_words` from the `tidytext` package and use an `anti_join` to remove all instances of that word.

```{r}
semeval_stance %>% 
  unnest_tokens(word, 'Tweet') %>%  # tokenize
  anti_join(stop_words)%>% # Remove stop words
  count(word, sort = TRUE) %>% # count by word
  arrange(desc(n)) %>% # Everything from this point on is just to graph
  head(20) %>% 
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_bar(stat= 'identity') + 
  coord_flip()
```

That looks better! Let's also take a look at the distribution of word counts by using a histogram on a log scale.

```{r}
semeval_stance %>% 
  unnest_tokens(word, 'Tweet') %>%  # tokenize
  anti_join(stop_words)%>% # Remove stop words
  count(word, sort = TRUE) %>%
  ggplot(aes(n)) +
  geom_histogram() +
  scale_x_log10()
```

It looks like we have a lot of words that only happen once, or otherwise very infrequently. We'll also remove some of the most infrequent words, as they are likely typos, or are so rare that they are not useful. The threshold is arbitrary and very much depends on the corpus itself. Here we will remove words that occur less than 2 times.

```{r}
semeval_stance %>% 
  unnest_tokens(word, 'Tweet') %>%  # tokenize
  anti_join(stop_words)%>% # Remove stop words
  count(word, sort = TRUE) %>% # count by word
  filter(n >= 2) %>%  # Remove words that occur less than 2 times
  ggplot(aes(n)) +
  geom_histogram() +
  scale_x_log10()
```

Now that we've explored the data a little bit and know what steps we should take to clean it up, we can create our features. From this exercise, we know we need to tokenize, remove stop words, and remove infrequent words. We can also take additional steps at this stage to clean up the data a bit more. For example, we might consider removing all numbers or digits. We can think about stemming (or lemmatization) in order to group similar words together under a single root (e.g., invent, invention, inventor). If we were to stem (this step is optional, and it's actually possible your models run better without stemming), it might look something like this:

```{r}
# NOTE: This uses the corpus package, which we did not bring in at the beginning
semeval_stance %>% 
  unnest_tokens(word, 'Tweet') %>%  # tokenize
  anti_join(stop_words)%>% # Remove stop words
  mutate(word = lemmatize_words(word) %>% unlist()) %>% # add stemming process
  count(word, sort = TRUE) %>% # count by word
  filter(n >= 10) %>%  # Remove words that occur less than 10 times
  arrange(desc(n)) %>% # Everything from this point on is just to graph
  head(20) %>% 
  ggplot(aes(x = reorder(word, n), y = n)) +
  geom_bar(stat= 'identity') + 
  coord_flip()
```

## Creating Features

To create the features for our machine learning model, we will take all of the words in our corpus and count the number of times that they appear in the tweet. We will end up with a sparse term-document frequency matrix, with the columns representing each word, and rows representing the tweet.

In this section, we will use the `tm` package to construct the feature matrix, as it offers a more convenient and intuitively interpretable approach.

Let's go over the steps to create the term-document frequency matrix:

1. lower all character
2. remove numbers, punctuation, and stop words
3. Stemming or lemmatization

Using `tm` package, we first need to create a corpus, which is a collection of text documents. We can then use the `tm_map` function to apply a series of text processing steps to the corpus.

```{r}
tweet_corpus = Corpus(VectorSource(semeval_stance$Tweet))
# text preprocessing
# Lower text
tweet_corpus = tm_map(tweet_corpus, content_transformer(tolower))
# Remove numbers, pubctuation, and stop words
tweet_corpus = tm_map(tweet_corpus, removeNumbers)
tweet_corpus = tm_map(tweet_corpus, removePunctuation)
tweet_corpus = tm_map(tweet_corpus, removeWords, c("the", "and", stopwords("english")))
# Lemmatization
tweet_corpus = tm_map(tweet_corpus, lemmatize_words)
```

After text preprocessing, we can create the term-document frequency matrix simply using the `DocumentTermMatrix` function.

```{r}
tweet_dtm = DocumentTermMatrix(tweet_corpus)
# Take a look at tweet_dtm
tweet_dtm
```
As we mentioned before, most of cases, the matrix will be sparse because most of the words will not appear in most of the tweets. We can use the `inspect` function to take a look at the first few rows and columns of the matrix.

```{r}
inspect(tweet_dtm)
```
One way to address the sparsity of the matrix is to remove infrequent words, as we demenstracted before. We can use the `removeSparseTerms` function to remove words that appear in less than a certain percentage of the tweets.

```{r}
tweet_dtm = removeSparseTerms(tweet_dtm, 0.99)
# Inspect the matrix again
inspect(tweet_dtm)
```
By specifying `0.99`, we are removing words that appear in less than 1% of the tweets. notably, we only have 105 unique terms left.

```{r}
# Convert the matrix to a data frame
tweet_dtm_df = as.data.frame(as.matrix(tweet_dtm))

# Add the ID, Stance variable to the data frame
tweet_dtm_df$ID = semeval_stance$ID
tweet_dtm_df$Stance = semeval_stance$Stance

```

The full data contains the ID variable (which we will use to make our train/test split in the next section), as well as our features (each word) and the label (True/False for whether the topic was on cell biology or not)

Here, you might consider doing a bit more feature engineering and data manipulation. For example, you might consider scaling the variables, in order to avoid the influence of more frequent words. You can try cleaning the text data a bit more, to remove certain words that might be stop words in this specific context. You can also consider adding additional variables, such as length of tweet in number of words.

## Train and Test Split

For simplicity, we'll consider a simple holdout sample. At the end, we show how to do cross validation using the `caret` package in R. The cross validation code will be very similar to this, except repeated for multiple combinations of training and testing data.

```{r, message=FALSE}
# 30% holdout sample
test <- tweet_dtm_df %>% sample_frac(.3)

# Rest in the training set
train <- tweet_dtm_df %>% anti_join(test, by = 'ID') %>% select(-ID)

# Remove ID after using to create train/test
# We don't want to use ID to run models!
test <- test %>% select(-ID)
```

# Fitting Models

Now, we can fit some machine learning models. We'll do some simple ones here: K-Nearest Neighbors and Support Vector Machine (SVM). You can also use Logistic Regression, or Naive Bayes, or Decision Trees, or any number of other, more complicated models, though we won't cover them here. If you are familiar with ensemble models, such as Random Forests, I'd suggest trying those out as well.

## First attempt at a model

Let's start with a K-Nearest Neighbors model. This simply checks the class of closest k neighbors, and takes a vote of them to predict what the class of the data point will be. We can fit this model using the `class` package.

```{r}
# Create separate training and testing features and labels objects
train_features <- train %>% select(-Stance)
test_features <- test %>% select(-Stance)

train_label <- train$Stance
test_label <- test$Stance

# Predicted values from K-NN, with K = 3
knnpred <- knn(train_features,test_features,train_label, k = 3)
```

The `knnpred` object has the predicted values for each of the `test_features` that we gave it. Let's take a look at what the predicted values are. We'll put the predicted values in a data frame with the actual values.

```{r}
pred_actual <- data.frame(predicted = knnpred, actual = test_label)
pred_actual %>% head()
```

Now that we have the predicted and actual values in one data frame, we can create a confusion matrix and evaluate how well our model is performing.

```{r}
pred_actual %>% table()
confusionMatrix(pred_actual %>% table())
```

Note that we don't actually see the words "precision" or "recall" here -- instead, we can find them by their alternate names: sensitivity (for recall) and positive predictive value (for precision). We can also use the `precision` and `recall` functions (also in the `caret` package). Note that we use `relevant` to specify which outcome we're trying to predict (similar to the `positive` argument above).

## Running a Support Vector Machine

With the training and testing datasets that we've created, running the actual tree model is actually quite simple. If you have used `R` for running linear models before, the format is very similar.

```{r}
svmfit <- svm(Stance ~ ., 
              data = train, 
              kernel = 'linear', 
              cost = 10)
```

Let's break down each of the arguments in this function. First, we specify the model, putting the label that we want to predict on the left side of the "~" and all the features we want to include on the right. We include arguments for the dataframe from which we're taking the data, and the kernel method we want to use. Then, we can use the `cost` argument to specify the regularization term.

We have stored the model in the `svmfit` object. Let's look at what the model gave us. We can use summary to look at the summary of the model.
```{r}
# You can try running the summary, but it will give a LOT of output
summary(svmfit)
```

### Evaluating the Model
Now that we have a model, we need to test it. We can get predictions using the `predict` function.

```{r}
pred <- predict(svmfit, test)
head(pred)
```

We can get the values of precision and recall using `confusionMatrix` function from the `caret` package. First, we create a table with the confusion matrix, then run the function with the table as the argument.

```{r}
# Construct the confusion matrix
pred_actual <- data.frame(predicted = pred, actual = test_label)
pred_actual %>% head()
confusionMatrix(pred_actual %>% table())
```

### Looping through models

We've shown an example of two different types of models: K-Nearest Neighbors and SVM. In the real world, we would want to try many different models with many different values for the tuning parameters (e.g., K for K-NN, cost or gamma for trees). Using a loop, we can automatically run through many different models very quickly. Below is an example of trying many different SVM models.

```{r}
# We will look at gamma values of 0.1, 1, 10, 100
gammas <- c(0.1, 1, 10, 100)

# We'll look at cost values of 0.1, 1, 10, 100, 1000
costs <- c(0.1, 1, 10, 100)

# We'll consider different kernel types
kernels <- c('linear', 'radial', 'polynomial')

# How many different models are we running?
nmods <- length(gammas)*length(costs)*length(kernels)

# We will store results in this data frame
results <- data.frame(gammas = rep(NA,nmods), 
                      costs = rep(NA, nmods),
                      kernel = rep(NA,nmods),
                      precision = rep(NA,nmods),
                      recall = rep(NA,nmods))

# The model number that we will iterate on (aka models run so far)
mod_num <- 1

# The loop
for(i in 1:length(gammas)){
  for(j in 1:length(costs)){
    for(k in 1:length(kernels)){
        g <- gammas[i]
        c <- costs[j]
        kn <- kernels[k]
        # Running the model

        svmfit <- svm(Stance ~ ., 
                      data = train, 
                      gamma = g,
                      cost = c,
                      kernel = kn)

        # Find the predictions
        pred <- predict(svmfit, test)

        # Attach scores to the test set
        # Then sort by descending order
        pred_actual <- data.frame(predicted = pred, actual = test_label)
        pred_table <- pred_actual %>% table()

        # Store results and we set AGAINST as referent category and turn the          results into a binary classification. 
        results[mod_num,] <- c(g, 
                               c,
                               kn,
                               precision(pred_table, relevant = 'AGAINST'), 
                               recall(pred_table, relevant = 'AGAINST'))
        # Increment the model number
        mod_num <- mod_num + 1
    }
  }
}

# All results are stored in the "results" dataframe
head(results)

# Best recall? Top 5 in descending order
results %>% arrange(desc(recall)) %>% head()

# Best precision? Top 5 in descending order
results %>% arrange(desc(precision)) %>% head()
```

You can also loop through many different values of K for K-NN, and also different tuning parameters for other types of models. For example, for Logistic Regression, you might consider an L1 or L2 penalty (Lasso or Ridge Regression), with different values for the penalty parameter.

## Simplifying the Process with Caret

We've already used the `caret` package for tools like the confusion matrix and precision/recall. However, we can also use it to do other validation methods more easily, such as k-fold cross validation. Here, we'll look at a quick example of using `caret` to find the training and test sets need to do 10-fold cross validation.

```{r}
# Create a list of 10 folds (each element has indices of the fold)
flds <- createFolds(tweet_dtm_df$ID, k = 10, list = TRUE, returnTrain = FALSE)
str(flds)
```

As you can see, we created a list of 10 vectors, each containing a fold. So, to do our 10-fold cross validation, we can take the first fold, and create our train and test sets from that, take the second fold and create test and train from that, and so on. For example, to create test and train using the first fold, we can use the following code:

```{r}
# Create train and test using fold 1 as test
stance_test01 <- tweet_dtm_df[flds$Fold01,]
stance_train01 <- tweet_dtm_df[-flds$Fold01,]
```

You can then use these train and test sets as we have above. You should put this into a loop, and store the value of metrics such as accuracy, precision, and recall, similar to what we've done in the loop of trying out different parameters.