### 2025-04-09 code/ml_llm/R/assign_3.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={DOGE's Downsizing, Can AI Read the Reddit Room?},
  pdfauthor={Kevin Linares; Felix Baez-Santiago; Aria Lu; Gloria Zhou},
  pdfkeywords={Reddit, Federal Government, DOGE},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{April 9, 2025}
\title{\bf DOGE's Downsizing, Can AI Read the Reddit Room?}
\author{
Kevin Linares\\
University of Maryland\\
and\\Felix Baez-Santiago\\
and\\Aria Lu\\
and\\Gloria Zhou\\
University of Michigan\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
We investigate public sentiment on Reddit regarding the Department of
Government Efficiency's federal workforce reduction by classifying 400
labeled comments (28\% favor, 18\% neutral, 54\% oppose) using
supervised and unsupervised Large Language Models (LLM). Supervised
models showed moderate success, particularly with ``oppose'' comments,
but struggled with ``favor'' and ``neutral'' stances. Similarly, LLMs
best identified ``oppose'' sentiment but exhibited low precision for
``favor'' and ``neutral.'' These findings highlight the challenges of
accurately gauging nuanced public opinion on government policy changes
using automated methods on social media data.
\end{abstract}

\noindent%
{\it Keywords:} Reddit, Federal Government, DOGE
\vfill

\newpage
\spacingset{1.9} % DON'T change the spacing!

\section{Introduction}\label{sec-intro}

The newly formed Department of Government Efficiency (DOGE) has reduced
the federal workforce by almost 280,000 employees, a central pledge of
the current administration. This action has generated significant
apprehension among federal workers in regards to mental health and job
security. To understand the impact of DOGE's actions on federal worker
perceptions of job security, we labeled 400 Reddit comments on topics
related to the current reduction in \emph{federal} workforce by DOGE as
whether the author favored, opposed, or had a neutral stance. We use
these labels to build supervised learning models to predict stance.
Additionally, we employ unsupervised large language models (LLM) to
detect the stance of these Reddit comments from text, to further explore
appropriate models for this current topic.

\section{Method}\label{sec-meth}

We collected a total of 12,553 Reddit comments in early March 2024 from
subreddits relevant to the topic of interest. These comments were
retrieved using keyword-based queries designed to capture posts of
discussions that would reflect users' stances on the federal workforce
reduction by DOGE. To prepare the data for analysis, a comprehensive
preprocessing pipline were conducted in Pytho 3. This process involved
removing HTML tags and web-URLs, stripping special characters and
tagging artifacts such as ``@'', and replacing emojis with textual
descriptions. Also, it entailed eliminating punctuations and converting
numerical values into their spelled-out forms. The text was standardized
to lowercase, lemmatized using the SpaCy library, and further refined by
removing stopwords per the Natural Language Toolkit (NLTK) lists.
Spelling corrections were performed to improve textual coherence, while
Duplicate comments were identified and removed to maintain data
integrity.

From this cleaned dataset, 400 uniqe comments without replacement were
randomly sampled for manual annotation. Four graudate students
independently coded and labeled each comment with one of three possible
stances:favor, opopose, or neutral toward DOGE's approach to federal
workforce reductions. As shown in Table 1, there were 28\% favor, 18\%
neutral, and 54\% oppose. These labeled comments were subsequently used
for training and evaluating both supervised and unsupervised
classification models.

\begin{longtable}[]{@{}lr@{}}
\caption{Distribution of labeled Reddit comment data.}\tabularnewline
\toprule\noalign{}
outcome & Percent \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
outcome & Percent \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 28 \\
neutral & 18 \\
oppose & 54 \\
\end{longtable}

The feature set used in model training and evluation consisted of two
primary categories: metadata features and textual features. Metadata
features included the Reddit comment's score, the subreddit in which it
appeared in, and the search term that retrieved it. As the variables for
score and upvotes were perfectly collinear in this dataset, only the
score was retained. To ensure all values were non-negative, the score
was adjusted by subtracting the dataset's minimum value from each score.
The categorical variables, subreddit and search term, were numerically
encoded using scikit-learn's label encoder.

Textual features were derived from each comment and represented using
two distinct approaches: unigram term frequency and term
freuqency-inverse document frequency (TF-IDF). The unigram used a
bag-of-words model to represent each word as a feature based on its
frequency in comment. The TF-IDF transformation built upon these raw
counts by accounting for term importance across the corpus. For both
unigram and TF-IDF methods, the total number of text-derived features
was 2,102. Combined with the three metadata features, each dataset
contained a total of 2,105 features. Two datasets were constructed for
classification: one combining the metadata with unigram features, and
another with metadat and TF-IDF features. Each dataset was divded into
80\% training data and 20\% test data. All supervised learning models
were implemented using Python 3, with scikit-learn and XGBoost as the
primary machine learning (ML) techniques. Classification was conducted
using a diverse set of algorithms, including Extra Trees Classifier,
Random Forest Classifier, Logistic Regression, Ridge Classifier,
K-Nearest Neightbors (KNN), Nearest Centroid, Multi-layer Perceptron
Classifier, Bernoulli Naive Bayes, Multino- mial Naive Bayes, Linear
Support Vector Classifier, Decision Tree Classifier, Extra Tree
Classifier, and the XGBoost Classifier.

All models were trained using default hyperparameter confiurations. The
number of neighbors for KNN model was set to three, and the mtry
parameter for Random Forrest was fixed at two, reflecting the relatively
small number of metadata features in those models. All experiments were
conducted on a personal computer without the use of a GPU

In addition to traditional supervised models, we employed two
instruction-tuned LLMs to perform zero- shot classification: Gemma
3.12B, developed by Google and based on the Gemini framework, and LLaMA
3.2B, developed by Meta AI. Gemma was deployed using the Great Lakes
High-Performance Computing (HPC) cluster, while LLaMA was deployed
locally on a GPU-enabled personal workstation using the rollama r
pacakge. These models were selected for their strong zero-shot
classification capabilities and ability to be run locally or on cluster
environments without reliance on commercial APIs.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Prompt}: & ``Is this comment in `favor', `neutral', or `oppose'
the reduction in federal workforce? Provide one word answer only! \\
\textbf{Task}: & ``You have assumed the role of a stakeholder that is
presented with a reddit comment from a likely federal worker related to
the current policies on reducing the federal workforce. Please determine
the author's stance on this topic, and only provide the answer.'' \\
\end{longtable}

The model performance was evaluated by the F1-score and recall. The
F1-score, a harmonic mean of precision and recall, serves as a balanced
indicator of a model's ability to correctly identify relevant instances
while minimizing both false positives and false negatives. Recall was
also independently assessed to evaluate the proportion of true postive
cases accurately identified by each model. While a high recall may
increase the risk of false positives, its combination with the F1-score
allows a more robust evaluation of model performance, especially for
addressing the challenges inherent in imbalanced classification tasks

\section{Results}\label{results}

Our machine learning models reveal a mix of performance results in
classifying the stance of Reddit comment into favor, neutral, oppose.
Both models exhibit low recall and F1-scores for the ``favor'' stance
(see Figure 1), yet the random forest model (recall .26: f1-score .28)
slightly outperforms KNN (recall .16: f1-score .19). The ``neutral''
class presented a significant challenge for the random forest with very
low scores (recall .05: f1-score .10), while KNN exhibited slightly
better performance (recall .21: f1-score .21). Both models showed the
strongest performance in identifying comments expressing opposition. The
random forest correctly identified 74 percent of all the actual Reddit
comments expressing opposition compared to KNN at 64 percent, meaning
that KNN missed a larger proportion of true ``oppose'' comments. The
f1-scores were similar for both models at 60 percent, this implies that
KNN may have a slightly higher precision score given that it had a lower
recall score than the random forest model. This implies that when KNN
predicts comments for the ``oppose'' class, it is as likely as the
random forest to be correct.

\includegraphics{assign_3_files/figure-pdf/unnamed-chunk-3-1.pdf}

The LLMs, showed poor performance in classifying the ``favor'' stance,
with very low f1-scores (llama .07: gemma .02) despite moderate recall
(llama .29: gemma .33), indicating low precision (see Figure 2). For the
``neutral'' class, both models showed modest results, with llama
achieving a slightly better F1-score (.23) compared to gemma (0.18),
although gemma had a higher recall (.50 vs .25), again suggesting lower
precision for gemma. Similar to the supervised models, both LLMs
performed best in classifying the ``oppose'' stance, with relatively
high and similar F1-scores (llama .67: gemma .70) and comparable recall
(llama .56: gemma .56), suggesting a better balance between precision
and recall for this category. Both models correctly identify 67 to 70
percent of all the actual Reddit comments expressing opposition.
Overall, both LLMs struggled with the ``favor'' and ``neutral'' classes
but demonstrated a stronger ability to identify opposing comments.\\

\includegraphics{assign_3_files/figure-pdf/unnamed-chunk-5-1.pdf}

Gemma and llama demonstrated only fair agreement in stance
classification (Cohen's Kappa = .24). The contingency table, Table 2
below, reveals highest agreement for ``oppose'' (.80), indicating some
consistency in identifying a strong negative stance. However, agreement
was substantially lower for ``neutral'' (.03) and ``favor'' (.01),
highlighting divergent interpretations of more nuanced language.
Off-diagonal values further illustrate these discrepancies, suggesting
fundamental differences in how the models process ambiguous cues and
establish classification boundaries, particularly for less extreme
stances.

\begin{longtable}[]{@{}lrrr@{}}
\caption{Aggreement between LLMs, Cohens Kappa, 0.24}\tabularnewline
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& favor & neutral & oppose \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
favor & 3 & 1 & 10 \\
neutral & 0 & 10 & 49 \\
oppose & 0 & 5 & 317 \\
\end{longtable}

\section{Conclusion}\label{conclusion}

We explored public sentiment on Reddit regarding the DOGE federal
workforce reduction using supervised learning (KNN, Random Forest) and
unsupervised LLMs ( gemma 3.12b and Llama 3.2 3B) for stance
classification. Our supervised learning models achieved moderate success
in this classification task, with the strongest performance in
identifying opposing viewpoints, which were also the most prevalent in
our labeled data. Moreover, both models struggled with the ``favor'' and
``neutral'' stances, suggesting limitations in our initial features to
capture these nuances. Our unsupervised learning models also showed the
best ability to classify ``oppose'' comments, yet exhibited significant
challenges with the ``favor'' and ``neutral'' categories, particularly
demonstrating low precision.

This study had limitations such as the small size of our labelled
dataset of 400 Reddit, imbalance in the stance categories, and few
features in the supervised approach. The zero-shot capabilities of the
LLMs are suitable for topic exploration but may require fine-tuning for
optimal performance on this specific domain, particularly with nuance
language used in ``favor'' comments. Future research should continue to
understand federal worker perceptions and improve stance classification
by expanding labeled dataset and incorporate more contextual text rather
than single comment text. Fine tuning LLMs on a larger corpus with
relevant comments could render more accuracy in the classification task.

\end{document}